---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Read data from csv.

```{r echo=TRUE}
activity <- read.csv("activity.csv")
summary(activity)
```

Converting the date from string to Date format
```{r echo=TRUE}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Historgrams for the number of steps table per day with mean and median
```{r echo=TRUE}
library(data.table)
library(ggplot2)
DT <- data.table(activity)
group_by_day <- DT[, sum(steps, na.rm=TRUE), by=date ]
setnames(group_by_day, "V1", "nosteps")
ggplot(group_by_day, aes(x = nosteps)) + geom_histogram(colour = "red")
mean(group_by_day$nosteps)
median(group_by_day$nosteps)
```



## What is the average daily activity pattern?

* Group the data by interval and average over all the days.
* Remove NAs when computing mean

```{r echo=TRUE}
group_by_time <- DT[, mean(steps, na.rm=TRUE), by=interval ]
setnames(group_by_time, "V1", "nosteps")
ggplot(group_by_time, aes(x = interval, y = nosteps)) + geom_line()
```

The Interval in which there was maximum number of steps
```{r echo=TRUE}
group_by_time[nosteps == max(group_by_time$nosteps), ][["interval"]]
```

## Imputing missing values

Total number of missing values
```{r echo=TRUE}
nrow(activity[is.na(activity$steps),])
```

**Stratergy for Imputing missing values**

For removing the missing values we find the mean steps for that interval over all the days and replace the missing value with that
```{r echo=TRUE}
group_by_time <- DT[, mean(steps, na.rm=TRUE), by=interval ]
setnames(group_by_time, "V1", "nosteps")
```

Creating the new dataframe from existing dataframe after removing missing values
```{r echo TRUE}
new_activity <- activity

for (n in 1: nrow(new_activity))
{
    if(is.na(new_activity$steps[n]))
    {
        new_activity$steps[n] <- group_by_time[group_by_time$interval == new_activity$interval[n],]$nosteps
    }
}

nrow(new_activity[is.na(new_activity$steps),])
```

Similar histograms as in Question 1
```{r echo=TRUE}
DT <- data.table(new_activity)
group_by_day <- DT[, sum(steps, na.rm=TRUE), by=date ]
setnames(group_by_day, "V1", "nosteps")
ggplot(group_by_day, aes(x = nosteps)) + geom_histogram(colour = "red")
mean(group_by_day$nosteps)
median(group_by_day$nosteps)
```
You can notice a few changes

  * Mean and Median are slightly higher
  * Lots of days when there are 0 steps decreased
  * The graph looks more like normal distribution now

## Are there differences in activity patterns between weekdays and weekends?

Creating a factor variable for weekday and weekend
```{r echo=TRUE}
activity$wend <- as.factor(ifelse(weekdays(activity$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday") )
```
```{r echo=TRUE}
DT <- data.table(activity)


group_by_time <- DT[, mean(steps,na.rm=TRUE), by=c("interval","wend")]
setnames(group_by_time, "V1", "nosteps")
ggplot(group_by_time, aes(x = interval, y = nosteps)) + geom_line(aes(color = wend)) + facet_grid(wend ~ .)
```

You can infer a new key differences with these plots

  * He starts to walk earlier during weekdays
  * But in weekends he ends up having much more steps overall
  * He stops walking later during weekends
